---
title: "Lab 9: Data Simulation Exploration"
author: "Eva Moylan"
date: "11/24/24"
format: 
  html:  
    embed-resources: true
    code-tools: true
editor: source
execute: 
  echo: true
code-fold: true
theme: sandstone
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(gt)
library(scales)

```

## Random Babies Simulation

Perhaps you have seen the [Random Babies applet](https://www.rossmanchance.com/applets/2021/randombabies/RandomBabies.html)? 
Suppose one night at a hospital four babies are born. The hospital is not very
organized and looses track of which baby belongs to each parent(s), so they 
decide to return the babies to parents at random. Here, we are interested in the
number of babies that are correctly returned to their respective parent(s).

**1. Simulate the distribution of the number of babies that are correctly returned. Use 10,000 simulations.**

::: callout-tip
Write a function to accomplish one simulation, then use `map_int()` to run
10,000 simulations. 

Keep in mind that your function needs to output a single number (not data frame) 
for it to be compatible with `map_int()`!
:::

```{r}
#| label: function-simulation-for-random-babies

randomBabies <- function(n = 4){ #This sets the default number of babies(n) to 4 if not otherwise specified
  
  baby_data <- tibble(baby = 1:n,
                      parent = sample(1:n,
                                      size = n,
                                      replace = FALSE)
  )
  
  baby_pairs <- baby_data |>
    mutate(correct_pair = if_else(parent == baby,
                                  TRUE,
                                  FALSE)) |>
    summarize(correct_pair = sum(correct_pair)) |>
    pull(correct_pair)
  
  return(baby_pairs) #return number of correct pairs
                  
}

results <- map_int(.x = 1:10000,
                   .f = ~ randomBabies(n = 4)
                   )

```

**2. Create a table displaying the proportion of simulations where 0, 1, 2, 3, and 4 babies were given to their correct parent(s).** Hint: A `pivot_wider()` will be helpful here!

::: callout-tip
The output of your `map_int()` is a vector, but to make a nice table (and plot) 
you need this to be a data frame! Luckily, the `enframe()` function does just 
that--it converts a vector to a data frame. 

You may find the following code helpful:

```{r}
#| eval: false

results <- enframe(results, 
        name = "simulation_number", 
        value = "Correct_Pairs")

```
:::

```{r}
#| label: table-for-random-babies

#Convert to dataframe
results <- enframe(results, 
        name = "simulation_number", 
        value = "Correct_Pairs")

#Make table
results |>
  count(Correct_Pairs) |>
  pivot_wider(names_from = Correct_Pairs,
              values_from = n) |>
  reframe(across(.cols = everything(),
                 .f = ~ (.x / 10000))
          ) |>
  mutate('3' = as.integer('0'),
         .after = '2') |> #Looks like there are no 3's, so add that to the table
  gt() |>
  tab_header(title = "Simulated Baby-Parent Matches",
             subtitle = "Proportion of simulations where 0, 1, 2, 3, and 4 babies were given to their correct parents."
  ) |>
  tab_footnote(footnote = "Note: 10,000 pairings simulated.") |>
  tab_style(
    style = cell_text(align = "center",
                      weight = "bold"), 
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(align = "center"), 
    locations = cells_body()
  ) |>
  fmt_percent(columns = everything())

#Changes made after submission: using fmt_percent to re-label proportions, and footnote added. 

```
**3. Now create a barplot showing the proportion of simulations where 0, 1, 2, 3, and 4 babies were given to their correct parent(s).** 

::: callout-tip
You may find the following code helpful:

```{r}
#| eval: false

geom_bar(mapping = aes(y = after_stat(count) / sum(after_stat(count))
                       )
         )
```
:::

```{r}
#| label: visualization-for-random-babies

results |>
  ggplot(aes(x = Correct_Pairs)) +
  geom_bar(mapping = aes(y = after_stat(count) / sum(after_stat(count)), 
           fill = after_stat(count) / sum(after_stat(count)))
           ) +
  theme_bw() +
  scale_fill_gradient(low = "tomato3",
                      high = "steelblue") +
  scale_y_continuous(labels = label_percent(suffix = "%"),
                     limits = c(0,
                                0.4)) +
  theme(legend.position = "none") +
  labs(y = NULL,
       x = "Number of Correct Pairings",
       title = "Simulated Proportion of 0 to 4 Correct Baby-Parent Pairings") 

#Changes made after submission: using label_percent to re-label the y axis, and changing wording of the title.

```

## Central Limit Theorem -- Optional & Somewhat Spicy

You have encountered the Central Limit Theorem in your previous statistics 
classes, whether or not is has been explicitly discussed. The Central Limit 
Theorem states that: 

> The sampling distribution of the mean will always be normally distributed, as
> long as the sample size is large enough, regardless of the underlying 
> distribution of the population. 

Remember back to your first statistics class when you had to check if the 
sample size was larger than 30 when testing if groups had different means? 
That's because of the Central Limit Theorem! Under certain conditions 
(e.g., sample size) the Central Limit Theorem ensures that the distribution 
of sample means will be approximately Normal, regardless of how skewed the 
underlying distribution of the population may be. 

A fundamental misunderstanding of the Central Limit Theorem is that it states 
that as a sample size gets larger, the population will be normally distributed. 
This is not the case, so let's do some exploring!  

**4. Write a function that simulates a specified number of sample means, for samples of size 100 drawn from a Chi-Squared distribution. Your function should allow the user to input:**

- **the number of means to simulate**
- **the degrees of freedom of the Chi-Squared distribution used to simulate data** 

I've provided some skeleton code to get you started. :) 

```{r}
#| label: simulate-means-function

simulate_means <- function(n, df){
  map_dbl(.x = 1:n, 
          .f = ~rchisq(n = 100, df = df) %>% mean()
          )
}

```

**5. Next, let's use the `crossing()` function to make a grid with inputs we want to pass into the `simulate_means()` function. Specifically, we want to explore the following values:**

-  **`n` = 10, 100, 1000, 10000**
-  **`df` = 10**

```{r}
#| label: create-grid

grid <- crossing(n = 10^seq(from = 1, 
                         to = 4,
                         by = 1), 
                 df = 10)
grid
```

**6. Now, use a `p_map()` to create a new column of simulated means (using the `simulate_means()` function), for every value in your `grid`.**

::: {.callout-tip}
You will want to use the `unnest()` function to extract the results of the
`p_map()` (stored in the `simulated_means` column). 
:::

```{r}
#| label: simulate-across-grid

all_simulations <- grid |> 
  mutate(simulated_means = pmap(.l = list(n = n, 
                                          df = df), 
                                .f = simulate_means)
         ) |> 
  unnest(cols = simulated_means) 

all_simulations
```

**7. Create a table of the means from each of the simulations (10, 100, 1000, and 10000).** 
Hint: Make sure your columns have descriptive names!

```{r}
#| label: table-of-simulated Means

all_simulations |>
  pivot_wider(id_cols = df,
              names_from = n,
              values_from = simulated_means,
              values_fn = mean) |>
  rename("Degrees of Freedom" = df,
         "Mean (n = 10)" = "10",
         "Mean (n = 100)" = "100",
         "Mean (n = 1000)" = "1000",
         "Mean (n = 10000)" = "10000") |>
  gt() |>
  tab_header(title = "Simulated Chi-Squared Means",
             subtitle = "Average simulated means across four simulation sizes."
  ) |>
  tab_style(
    style = cell_text(align = "center",
                      weight = "bold"), 
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(align = "center"), 
    locations = cells_body()
  )

#Change made after submission: Subtitle changed from 'sample sizes' to 'simulation sizes'

```

**8. Create a plot showing the distribution of simulated means from each of the simulations. Each simulation (10, 100, 1000, and 10000) should be its own facet!**
Hint: Make sure your facets have descriptive names! You might also want to free
the y-axis of the plots, since there are substantial differences in the sample
sizes between the simulations. 

**For extra pizzaz, add a vertical line for true mean (for a Chi-Square the mean is the degrees of freedom).**

```{r}
#| label: plot-of-simulated Means
#| message: false

all_simulations |>
  mutate(across(.cols = n:df,
                .fns = ~as.character(.x)),
         n = fct_recode(n,
                        `n = 10` = "10",
                        `n = 100` = "100",
                        `n = 1000` = "1000",
                        `n = 10000` = "10000"),
         df = fct_recode(df,
                         `Degrees of Freedom = 10` = "10")) |>
  ggplot(mapping = aes(x = simulated_means)) +
  geom_histogram(color = "white") +
  facet_grid(n ~ df,
             scales = "free_y") +
  geom_vline(xintercept = 10, 
             color = "tomato") +
  theme(axis.text.y = element_text(size = 7),
        aspect.ratio = 0.2) +
  labs(y = NULL,
       x = "Simulated Means",
       subtitle = "Frequency of Means",
       title = "Simulated Means for Four Different Simulation Sizes")

#Change made after submission: Subtitle changed from 'sample sizes' to 'simulation sizes'

```

## Challenge 9

Instructions for the challenge can be found on the course website or through 
the link in Canvas! 

